%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% Databae Object and DDL
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%========= SECTION SECTION ===================================================

\section{DDL Propagation to Other Nodes}
  
  As mentioned in the architecture section~\ref{sec:ddlPropagation} at page~\pageref{sec:ddlPropagation},
  \XC{} propagates DDL execution to other
  node except for node management statements:
  \texttt{CREATE NODE}, \texttt{ALTER NODE}, \texttt{DROP NODE}, 
  \texttt{CREATE NODE GROUP}, and \texttt{DROP NODE GROUP}.
  
  In \PG, functions handling DDL statements start with those in
  \file{backend/tcop/utility.c}.
  Parsed DDL statements are fist handled by functions \file{ProcessUtility()}.
  If no hook is defined, then it is passed to \file{standard_ProcessUtility()}.
  If target object supports event triggers, then they are passed to
  \file{ProcessUtilitySlow()}.
  
  DDL propagation to other nodes is implemented in each DDL execution code.
  
  The following list shows an example for handling \texttt{CREATE TABLESPACE} statement
  in \file{standard_ProcessUtility()}.

  % Source list of the example: DDL propagation to other node
  \lstset{tabsize=4, xleftmargin=20pt, basicstyle=\ttfamily\scriptsize, breaklines=true}
  \begin{lstlisting}[frame=single, tabsize=4, language=C]
        case T_CreateTableSpaceStmt:
#ifdef PGXC
            if (IS_PGXC_COORDINATOR && !IsConnFromCoord())
#endif
            /* no event triggers for global objects */
            PreventTransactionChain(isTopLevel, "CREATE TABLESPACE");
            CreateTableSpace((CreateTableSpaceStmt *) parsetree);
#ifdef PGXC
            if (IS_PGXC_COORDINATOR && !IsConnFromCoord())
                ExecUtilityWithMessage(queryString, sentToRemote, false);
#endif
            break;
  \end{lstlisting}
  % End program listing -------------------------------------------------------

  The first block of the directive ``\file{# ifdef PGXC}'' tests if the current node is coordinator and
  the DDL is not from another coordinator.
  If so, then this is the root and has to check transaction chain.
  
  The second block of the directive ``\file{# ifdef PGXC}'' again tests if it the current node is a coordinator and
  the DDL is not from another coordinator.
  If so, this node has to take care of DDL propagation.
  Otherwise, it just concentrate on local handling.
  
  DDL propagation is handled in the same manner for other DDSs as well.


%========= SECTION SECTION ===================================================

\section{Additional Error Handling}

  If all the DDL can be handled within a transaction block and can handle abort correctly in vanilla \PG,
  \XC{} can use implicit 2PC to handle errors in other nodes.
  Unfortunately, some DDL such as \texttt{CRETE TABLESPACE} cannot be issued inside transaction block.
  \XC{} has to propagate such DDL too and has to complete the statement in atomic way.
  In other words, \XC{} has to guarantee that the DDL cannot be successful only partly in some nodes.
  \XC{} needs separate feature to maintain cluster-wide
  data integrity in such case, without using 2PC.
  
  To make this cleanup work, \XC{} defines internal structure \file{dbcleanup_info} of the type
  \file{abort_callback_type} inside \file{execRemote.c}.
  
  \file{set_dbcleanup_callback()} function registers cleanup function and cleanup data specific to each
  DDL which run only outside a transaction block.
  If this is registered, then \file{AtEOXact_DBCleanup()} will invoke it at the end of a transaction.


%========= SECTION SECTION ===================================================

\section{Additional Functions to handle DDL}

  \file{utility.c} is an entry point of most of the DDL handlers.
  To propagate DDL to other nodes, \XC{} implements several utility functions in this module.


%------- Func ----------------------------------------------------------

\FUNC{IsStmtAllowedInLockedMode()}

  Determines if a given statement can run within a transaction block.
  It is used to determine if dedicated error handling is needed.


%------- Func -----------------------------------------------------

\FUNC{ExecUtilityWithMessage()}

  This function performs a statement in a remote node within a transaction block.
  This handles error by attaching failed node name and by rethrowing it.


%------- Func -----------------------------------------------------

\FUNC{ExecUtilityStmtOnNodes()}

  This function executes a utility statement on nodes, including coordinators.


%------- Func -----------------------------------------------------

\FUNC{ExecUtilityFindNodes()}

  Determines the list of nodes to launch query on.


%------- Func -----------------------------------------------------

\FUNC{ExecUtilityFindNodesRelkind()}

  Determines which node a statement should be executed on the given relation.


%------- Func -----------------------------------------------------

\FUNC{GetNodesForCommentUtility()}

  Returns Object ID of object commented.


%------- Func -----------------------------------------------------

\FUNC{GetNodesForRulesUtility()}

  Gets the nodes to execute the given RULE related to a utility statement.


%------- Func -----------------------------------------------------

\FUNC{DropStmtPreTreatment()}

  Performs a pre-treatment of \texttt{DROP} statement on a remote coordinator.


%========= SECTION SECTION ===================================================

\section{\label{sec:tablespace}Tablespace}

\input{tablespace.tex}


%========= SECTION SECTION ===================================================

\section{\label{sec:mview}Materialized View}

\input{matview.tex}


%========= SECTION SECTION ===================================================

\section{\label{sec:updatableView}Automatic Updatable View}

\input{upd_views.tex}


%========= SECTION SECTION ===================================================

\section{\label{sec:trigger}Trigger}

\input{trigger.tex}


%========= SECTION SECTION ===================================================

\section{\label{sec:eventTrigger}Event Trigger}

\input{event_trigger.tex}


%========= SECTION SECTION ===================================================

\section{\label{sec:tempObject}Temporary Objects}

  The change has been made to allow temporary object usage in explicit 2PC transactions.
  
  The background that temporary object is not allowed in 2PC is that \textbf{PREPARED} transaction survives
  the session, while temporary objects do not.
  
  In contrary, implicit 2PCs do not survive the session.
  Even with crashes, \file{pgxc_clean} cleans up implicit 2PC transactions so that they do not
  survive.
  It is safe to allow temporary objects to be used in implicit 2PCs.
  
  Changes are done in \file{CommitTransaction()} in \file{xact.c}.
  
  The patch is as follows:

  % Patch to support temporary objects.
  \lstset{tabsize=4, xleftmargin=20pt, basicstyle=\ttfamily\scriptsize, breaklines=true}
  \begin{lstlisting}[frame=single, tabsize=4, language=C]
--- a/src/backend/access/transam/xact.c
+++ b/src/backend/access/transam/xact.c
@@ -2117,6 +2117,9 @@ CommitTransaction(void)
 {
	TransactionState s = CurrentTransactionState;
	TransactionId latestXid;
+#ifdef PGXC
+       bool            isImplicit = !(s->blockState == TBLOCK_PREPARE);
+#endif
							  
		ShowTransactionState("CommitTransaction");

@@ -2161,7 +2164,7 @@ CommitTransaction(void)
				*/
				if (IsOnCommitActions() || ExecIsTempObjectIncluded())
				{
-                       if (!EnforceTwoPhaseCommit)
+                       if (!EnforceTwoPhaseCommit || isImplicit)
							ExecSetTempObjectIncluded();
						else
							ereport(ERROR,
@@ -2655,7 +2658,11 @@ PrepareTransaction(void)
			* cases, such as a temp table created and dropped all within the
			* transaction.  That seems to require much more bookkeeping though.
			*/
+#ifdef PGXC
+       if (MyXactAccessedTempRel && !isImplicit)
+#else
		if (MyXactAccessedTempRel)
+#endif
			ereport(ERROR,
					(errcode(ERRCODE_FEATURE_NOT_SUPPORTED),
					 errmsg("cannot PREPARE a transaction that has operated on temporary tables")));
  \end{lstlisting}
